The core transcriptome of mammalian placentas and the divergence of expression with placental shape

### 2017-03-08 manuscript/manuscript.Rnw
describe how the fetal villous portion was separated

Fetal tissue was collected from the placenta of nine mammalian
species (\human [human], \bonobo [bonobo], \smonkey [spider monkey], \dog [domestic dog], \cow
[domestic cow], \elephant [African savanna elephant], \opossum [gray
short-tailed opossum], \galili [blind mole rat from Upper Galilee],
\carmeli [blind mole rat from northern Israel]
\cite{Nevo2001:adaptiveradiationofblindsubterraneanmole};
n=1 for all collected species). Information
on specimen ID, number of specimens, and collector is shown for each sample in Supplemental
\cref{tab:srr_numbers}. In the case of \bonobo, \cow, and \elephant,
the placenta was collected at term upon expulsion from the mother
following delivery. Placental tissue from \human was collected upon
delivery via a cesarean section. Near-term placental tissue from \dog
was collected upon the death of the mother and fetus from a car accident.
Placental tissue from \opossum was sampled at fetal stage 32, roughly
one day prior to birth. Only the fetal portion of the placenta was
sampled at the maternal-fetal interface (i.e. villous portion for
\human, \bonobo, \dog, and \elephant; cotyledon for \cow; diffuse
placenta for \opossum). All tissues were collected and stored in
RNAlater (Life Technologies, Carlsbad, CA) until RNA isolation could
be performed. Tissues were homogenized in Trizol, and total RNA was
isolated following the manufacturer's instructions. RNA was further
purified using RNeasy Midi Kit (Qiagen, Valencia, CA), in conjunction
with the RNase-Free DNase Set (Qiagen, Valencia, CA), following the
manufacturer’s protocol to remove any contaminating DNA from the
sample. Total RNA was then evaluated on a Nanodrop 1000 (Thermo
Scientific, Wilmington, DE) and an Agilent Bioanalyzer 2100 (Santa
Clara, CA) to confirm concentration and quality of the isolation. RNA
quality was assessed using A260/A280 and A260/A230 absorbance values.

>>>>

Fetal tissue was collected from the placenta of nine mammalian species
(\human [human], \bonobo [bonobo], \smonkey [spider monkey], \dog
[domestic dog], \cow [domestic cow], \elephant [African savanna
elephant], \opossum [gray short-tailed opossum], \galili [blind mole
rat from Upper Galilee], \carmeli [blind mole rat from northern
Israel] \cite{Nevo2001:adaptiveradiationofblindsubterraneanmole}; n=1
for all collected species). Information on specimen ID, number of
specimens, and collector is shown for each sample in Supplemental
\cref{tab:srr_numbers}. In the case of \bonobo, \cow, and \elephant,
the placenta was collected at term upon expulsion from the mother
following delivery. Placental tissue from \human was collected upon
delivery via a cesarean section. Near-term placental tissue from \dog
was collected upon the death of the mother and fetus from a car
accident. Placental tissue from \opossum was sampled at fetal stage
32, roughly one day prior to birth. Only the fetal villous portion of
the placenta was sampled at the maternal-fetal interface (i.e. villous
portion for \human, \bonobo, \dog, and \elephant; cotyledon for \cow;
diffuse placenta for \opossum). The placenta membranes were separated
from the fetal villous portion using forceps and a scapel, and the
resulting fetal villous portion was washed twice in PBS. All tissues
were collected and stored in RNAlater (Life Technologies, Carlsbad,
CA) until RNA isolation could be performed. Tissues were homogenized
in Trizol, and total RNA was isolated following the manufacturer's
instructions. RNA was further purified using RNeasy Midi Kit (Qiagen,
Valencia, CA), in conjunction with the RNase-Free DNase Set (Qiagen,
Valencia, CA), following the manufacturer’s protocol to remove any
contaminating DNA from the sample. Total RNA was then evaluated on a
Nanodrop 1000 (Thermo Scientific, Wilmington, DE) and an Agilent
Bioanalyzer 2100 (Santa Clara, CA) to confirm concentration and
quality of the isolation. RNA quality was assessed using A260/A280 and
A260/A230 absorbance values.
